\documentclass[journal]{IEEEtran}


\title{Collaborative Learning with Dirichlet Process Clustering for Rapid Online Adaptation in Robotics}
\author{Runjun Mao \\ August 12, 2023 \\ Supervised by Antoine Cully}
%\date{August 12, 2023}


\usepackage{bm}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\renewcommand{\algorithmiccomment}[1]{\hfill $\triangleright$ #1}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\hypersetup{hidelinks}



\begin{document}
\maketitle

\begin{abstract}
\noindent
Robots in deployment can face unforeseeable distortion in their motion due to unseen environment or hardware failures. 
Model-based online adaptation learns this distortion from the interactions with the current environment and corrects its motions accordingly. However, this can be very challenging as it is hard to select the most suitable model and its hyper-parameters for the current dynamics with little prior knowledge.
If we could leverage the plentiful historical real-world interactions, we may build statistical models for the common types of dynamics faced by the robot, hence providing guidance for adaptations. 
This paper developed a method to enable fast online adaptation for the robot by grouping up historical  interaction data into clusters. 
The historical data can be collected from several robots of the same design performing different tasks in different environments, hence allowing learning in a collaborative manner. 
The method takes an non-parametric approach of novelty by modelling the data distribution with infinite mixture of Gaussian Processes and conducting clustering using Dirichlet Processes. 
This achieves unbiased training while ensures high efficiency in terms of both data and computation during deployment. 
The experiment is conducted on a simulated robot to compare with the previous work for online adaptation. 
The method accurately groups up the collaboratively collected data without any knowledge of their sources, and it offers a significant improvement over the baseline in both the efficiency of learning the distortion and the capability of adaptation. 

\end{abstract}


\input{introduction.tex}

\input{background.tex}

\input{related_works.tex}

\input{methodology.tex}

\input{experiment.tex}


\input{results_and_analysis.tex}
%To formulate the problem clearly, we are aiming to leverage the collaboratively collected real-world interactions to assist a robot deployed in an unknown environment to quickly adapt and complete its task.
%We are provided with some episodes (each task-solving process corresponds to an episode of interaction data) of real-world interactions collected by several robots deployed in an unknown distribution of environments.
%The most intuitive idea is to label these historical data with their sources (like collected from grass terrain, collected with left leg broken, etc.) and build a simple database in the cloud server. 
%During the adaptation, the robot examines the current environment and uploads the result to the database server. The server then identifies the most similar environment recorded in the database and let the robot download the corresponding data to assist the its adaptation.
%However, accurately measuring and describing the environment can be expensive and difficult. It is also hard to determine whether a previously encountered environment shares similar dynamics with the current one even if we have accurate descriptions for them.
%In the case we do know that some historical data come from exactly the same environment as the current one, these data might be insufficient in quantity to assist the current task.
%Hence, it is reasonable to group the episodes resulting from similar dynamics into clusters, ensuring having enough data for analysis.
%We will assume that we don’t have access to any information of the environment and aim to conduct clustering only according to the dynamics. 



\section{Conclusion}





\bibliographystyle{unsrt}
\bibliography{reference}

\end{document}